---
title: "GoogleData"
author: "John Bernau"
date: "10/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overview  
RQ: How do mass shooting events influence search traffic?  
DATA: Daily search traffic for terms like "gun+control" from 2004-2018. Extracted 14-day event windows around each shooting and appended shooting characteristics (race of shooter, victims, venue, etc).    

___  

<img src = "/private/sample2.jpg">
<img src = "/private/gt_race.jpg">
<img src = "/private/gt_state.jpg">

___  

Sample dataset for one term. Looks like panel data...  
<img src = "/private/sample.png">

```{r, message = FALSE}
load("/Users/johnbernau/Box Sync/1. Desktop/3. Publications/Google_Search_Data/Analysis/9.19.18 session/gc_cleaned.RData")
require(tidyverse)
require(broom)
require(plm)
```


```{r}
t5 <- filter(t5, days_after < 15)
names(t5)
```


```{r, warning = FALSE}
ggplot(t5, aes(n_weapons2)) + geom_histogram(stat = "count")
ggplot(t5, aes(highest_weapon)) + geom_histogram(stat = "count")
```


Save as panel data format
```{r}
pt <- pdata.frame(t5, index = c("id", "days_after"))
```



Gives variance of DV? 69% variance between, 3% variance within?
```{r}
summary(pt$hits_final)
```


## Pooling OLS Model
These results would work best with our theory...  
```{r}
pooling <- plm(hits_final ~ total_victims + venue + race + n_weapons2 + highest_weapon + days_after, data=pt, model= "pooling")
summary(pooling)
```


## Between model
```{r}
between <- plm(hits_final ~ total_victims + venue + race + n_weapons2 + highest_weapon + days_after, data=pt, model= "between")
summary(between)
```


## Fixed effect  
(Also known as "within"" estimator. Variables that don't vary within clusters are dropped.)  
```{r}
fixed <- plm(hits_final ~ total_victims + venue + race + n_weapons2 + highest_weapon + days_after, data=pt, model= "within")
summary(fixed)
```


(OLS with dummies for group is identical to fixed effects)
```{r, eval = FALSE}
ols_model <- lm(hits_final ~ days_after + id, data=pt)
summary(ols_model)
```


## Random effect 
Splits the difference between regular pooled OLS (?) and fixed effects.  
```{r}
random <- plm(hits_final ~ total_victims + venue + race + n_weapons2 + highest_weapon + days_after, data = pt, model = "random")
summary(random)
```

## Questions:  
* How do you know which model to use?  
* First-differences estimator?  
* Is model controlling for time, or do you have to add that to the model explicitly?  
* Interaction effects? Namely, shooting characteristics and time.  
* Random slopes?  Panel models vs multi-level models?  
* Dynamic panel models? Autocorrelation of DV?  

___  

## Postestimation:  
LM test for random effects versus OLS. If significant, use random effects over OLS. 
```{r}
plmtest(pooling)
```


LM test for fixed effects versus OLS. If significant, use fixed effects over OLS. 
```{r}
pFtest(fixed, pooling)
```


Hausman test for fixed versus random effects model. If significant and "one model is inconsistent" use fixed effects over random. 
```{r}
phtest(random, fixed)
```



